\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-04-30 paper.qmd
notation update

The 2D hexagon grid is defined by the number of bins in each direction ($b_1, b_2$), and the lower left position where the grid starts ($s_1, s_2$). The values of $s_i$ need to be below their respective minimum variable values, and could be a full bin lower, to allow a buffer corresponding to a full hexagon width ($a_1$) and height ($a_2$) around the data. The values of $b_i$ are variables to be computed that define the reduction in size of the data ($n$ to $m$).

>>>>

The 2D hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, as given by the centroids of each hexagon $(c_1, c_2)$ and the lower left position where the grid starts at $(s_1, s_2)$, which correspond to the lowest left centroid. The values of $s_i$ need to be below their respective minimum variable values, and could be a full bin lower, to allow a buffer corresponding to a full hexagon width ($a_1$) and height ($a_2$) around the data. The values of $b_i$ are variables to be computed that define the reduction in size of the data ($n$ to $m$).

### 2024-04-30 paper.tex
Corrected vs formular
Changed vs notation
notation update

\begin{tabular}{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{12cm}}
\toprule
\textbf{Notation} & \textbf{Description}\\
\midrule
$n, p, k, m$ & number of observations, variables, embedding dimension, number of non-empty bins, respectively\\
$\mathbfit{X}, \mathbfit{x}$ & $p$-dimensional data (population, sample)\\
$\mathbfit{y}$ & $k$-dimensional layout\\
$P$ & orthonormal basis, generating a $d\text{-}dimensional$ linear projection of $p$-dimensional data\\
$T$ & true  model\\
\addlinespace
$g$ & functional mapping from \pD{} to \kD{}, especially as prescribed by NLDR\\
$\mathbfit{\theta}$ & (Hyper-) parameters for NLDR method\\
$y_\text{max}$ & maximum value of the scaled second embedding component\\
$r$ & ranges of the embedding components\\
$b$ & bins\\
\addlinespace
$a_r$ & aspect ratio of the 2D layout before scaling\\
$s$ & hexagonal size (defined as the radius of the outer circle surrounding the hexagon)\\
$(s_1, \ s_2)$ & starting coordinates of the hexagonal grid\\
$q_1$, $q_2$ & buffer amount along the x and y axes\\
$h_s$, $v_s$ & horizontal and vertical distance between adjacent hexagon bin centroids\\
\addlinespace
$d_x$, $d_y$ & horizontal spacing and adjusted vertical spacing for hexagonal coordinates\\
$b_1$, $b_2$ & number of hexagon bins along the x and y axes\\
$b$, $b'$ & total and non-empty hexagon bins in the grid\\
$n_k$ & number of observations within the $k^{th}$ hexagon\\
$(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of the $k^{th}$ hexagon\\
\addlinespace
$C^{(j)}$ & $j$-dimensional bin centers\\
\bottomrule
\end{tabular}
\endgroup{}

@@

The 2D hexagon grid is defined by the number of bins in each direction
(\(b_1, b_2\)), and the lower left position where the grid starts
(\(s_1, s_2\)). The values of \(s_i\) need to be below their respective
minimum variable values, and could be a full bin lower, to allow a
buffer corresponding to a full hexagon width (\(a_1\)) and height
(\(a_2\)) around the data. The values of \(b_i\) are variables to be
computed that define the reduction in size of the data (\(n\) to \(m\)).

The value for \(b_2\) is computed by fixing \(b_1\). Considering the
lower bound of the NLDR, \(a_1 > 2 \times s_1\) and
\(a_1 > \frac{1-s_1}{b_1 -1}\). The minimum \(a_1\) that satisfies the
constraints will be selected. For regular hexagons,
\(a_2 = \frac{2}{\sqrt(3)}a_1\). Therefore, \(a_2\) is computed for the
minimum \(a_1\). According to the upper bound of the NLDR,
\(b_2 = \left\lceil\frac{r_2 - s_2}{vs} + 1\right\rceil\), where
\(vs = \frac{a_2}{2} + \sqrt{{(\frac{a_2}{2})}^2 + {(\frac{a_1}{2})}^2}\).

>>>>

\begin{tabular}{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{12cm}}
\toprule
\textbf{Notation} & \textbf{Description}\\
\midrule
$n, p, k, m$ & number of observations, variables, embedding dimension, number of non-empty bins, respectively\\
$\mathbfit{X}, \mathbfit{x}$ & $p$-dimensional data (population, sample)\\
$\mathbfit{y}$ & $k$-dimensional layout\\
$P$ & orthonormal basis, generating a $d\text{-}dimensional$ linear projection of $p$-dimensional data\\
$T$ & true  model\\
\addlinespace
$g$ & functional mapping from \pD{} to \kD{}, especially as prescribed by NLDR\\
$\mathbfit{\theta}$ & (Hyper-) parameters for NLDR method\\
$r$ & ranges of the embedding components\\
$C^{(j)}$ & $j$-dimensional bin centers\\
$b$ & number of bins in each direction\\
\addlinespace
$a$ & binwidths, distance between centroids in each direction\\
$(s_1, \ s_2)$ & starting coordinates of the hexagonal grid\\
$q$ & buffer to ensure hexgrid covers data, proportion of data range, 0-1\\
$b$, $b'$ & total and non-empty hexagon bins in the grid\\
$n_k$ & number of observations within the $k^{th}$ hexagon\\
\bottomrule
\end{tabular}
\endgroup{}

@@

The 2D hexagon grid is defined by the number of bins in each direction
\((b_1, b_2)\), as given by the centroids of each hexagon \((c_1, c_2)\)
and the lower left position where the grid starts at \((s_1, s_2)\),
which correspond to the lowest left centroid. The values of \(s_i\) need
to be below their respective minimum variable values, and could be a
full bin lower, to allow a buffer corresponding to a full hexagon width
(\(a_1\)) and height (\(a_2\)) around the data. The values of \(b_i\)
are variables to be computed that define the reduction in size of the
data (\(n\) to \(m\)).

The value for \(b_2\) is computed by fixing \(b_1\). Considering the
lower bound of the NLDR, \(a_1 > 2 \times s_1\) and
\(a_1 > \frac{1-s_1}{b_1 -1}\). The minimum \(a_1\) that satisfies the
constraints will be selected. For regular hexagons,
\(a_2 = \frac{2}{\sqrt(3)}a_1\). Therefore, \(a_2\) is computed for the
minimum \(a_1\). According to the upper bound of the NLDR,
\(b_2 = \left\lceil\frac{r_2 - s_2}{v_2} + 1\right\rceil\), where
\(v_2 = \frac{a_2}{2} + \sqrt{{(\frac{a_2}{2})}^2 - {(\frac{a_1}{2})}^2}\).
